Response to the defense feedback

### 2021-06-15 Chapter-5/Chapter-5.tex
Updates from Overleaf

\chapter{Precise Analysis on Multiple-trace Attacks}\label{chapter5}
\section{Introduction}
Side-channel attacks allow attackers to infer sensitive information based on the non-functional behaviors of the computer system. Examples of these non-functional behaviors include acoustics, CPU usages, timing, EM signals, etc~\cite{agrawal2002side,hund2013practical,halevi2015keyboard,batina2019csi}. In summary, if a program has different execution behaviors when it processes various keys, an attacker can infer nontrivial information by observing secret-dependent behaviors. While separate side-channel attacks may have different forms and patterns, a large portion of them share the same fundamental reasons. That is, the memory access pattern is dependent on the original sensitive input.

@@

Even a minor leakage can be severe for cryptography libraries because
the leakage can reduce the encryption algorithm's strength. However,
some attacks also exploit non-cryptography libraries and applications
such as machine learning
applications~\cite{yan2020cache,hong2018security}, graphic
libraries~\cite{wang2019unveiling}, spell checking
tools~\cite{xu2015controlled}, etc. Unlike side-channel attacks on
cryptography libraries that exploit one vulnerability at a time,
attacks on non-cryptography rely on multiple side-channel leakage
sites to retrieve more information. A program  usually has more than two
leakage sites. However, no existing tools that can estimate the total
effect of multiple leakage sites. For real-world applications with
thousands of lines, attacker can exploit more than one leakage sites
to retrieve more information. Attackers can guess the values of many temporary
values. And these temporary values contain some knowledge of the
original secret buffer.

Second, many side-channel leakage detection tools rely on
techniques such as symbolic executions, taint analysis, and abstract interpretations~\cite{203878,182946,Brotzman19Casym,236338}. Due
to the limitations of existing binary analysis techniques, it is hard to apply existing methods to analyze side-channels in non-cryptography
libraries. For many other applications like machine learning, media,
and graphic libraries built on the top architecture-dependent
libraries (e.g., Intel MKL~\cite{wang2014intel}), these tools are not able to
handle them. Moreover, some binary analysis frameworks (e.g.,
Angr~\cite{shoshitaishvili2016state}) are limited in analyzing
floating-point instructions. However, the functionality of these
applications heavily depends on floating-point calculations.

Third, while many existing works can detect side-channel leakages, few
of them can assess the severity levels of these
vulnerabilities~\cite{182946,203878,182946,236338,217537,langley2010ctgrind,xiao2017stacco,wang2019time,Wichelmann:2018:MFF:3274694.3274741,bao2021abacus,236334,wang2021data,sung2018canal}.
Some existing tools~\cite{182946,chattopadhyay2019quantifying, biondi2018scalable,5207642,pasareanu2016multi,AskarovC12,phan2017synthesis} quantify the information leakage by giving an upper
bound estimation of each leakage site.  This is helpful to justify that the software is secure if the reported leakage is zero. However,
these tools cannot report to software developers how severe each leakage site would be because of the over-approximation they apply when they estimate the amount of leakage.  Often the over-approximated estimate is very far away from the amount of real leakage.
For example, CacheAudit~\cite{182946} estimates that a 128-bit
AES encryption can leak more than 128 bits. As a result, even these
tools report that some leakages can leak lots of information, it does not mean the program has truly severe side-channel vulnerabilities.

To solve the above problem, we propose a method to estimate the effect of side-channel leakages automatically. We treat the side-channel attack as a communication system and use the channel capacity to quantify the amount of the information flow from the original secret to an attacker's observation. The method also allows us to combine multiple leakages sites to retrieve the information. The attack can be seen as a process that reduces the search space of the original sensitive inputs. We use the side-channel vulnerability to divide the input space. If the patterns can uniquely distinguish the input space, then we think the information is leaked totally. However, in real cases, many of these side-channel leakage patterns are not sufficient to distinguish each leakage site but can still discern some secret inputs. In these cases, we call them partial leakages.

@@

  \item We propose a method that automatically detects and analyzes the
  effect of each side-channel leakage site. Our analysis combines
  information from both the source code and the run-time execution,
  which makes the method more effective in finding the leakages. The
  technique can combine multiple leakage sites to retrieve more
  information as well.

@@

\section{Background and Threat Model}
In this chapter, we treat the side-channel attack as a communication system between two parties.
In the section, we give an overview of the background knowledge of information theory and present the threat model of the chapter.

\subsection{Channel Capacity}
Given a discrete random variable $K$,
$k$ is one of the possible
values occurring with the probability $p(k)$. The Shannon entropy $H(K)$ is defined as
\begin{equation}\label{chapter5:eq1}
  H(K) = - \sum_{k {\in} K}p(k)\log_2p(k)
\end{equation}

@@

\begin{equation}\label{chapter5:eq2}
  C(K;O) = \max_{p(x)} I(K;O)
\end{equation}
Here $I(K;O)$ is the mutual information between $K$ and $O$.
\begin{equation} \label{chapter5:eq3}
  I(K;O) = \sum_{k {\in} K}{\sum_{o {\in} O}{p(k, o)\log_2\frac{p(k, o)}{p(k)p(o)}}}
\end{equation}

@@

\subsubsection{Noiseless Lossless Channel}
An information channel is noiseless and lossless if every $k$ corresponds to exactly one unique $o$. The channel capacity of a noiseless lossless channel equals to
\begin{equation} \label{chapter5:eq4}
  C(K;O) = \max_{p(k)} I(K;O) = \max_{p(k)} H|O| =\max_{p(k)} H|K|
\end{equation}
$|K|$ represents the number of possible variables in $K$.

Figure~\ref{fig:channel}(a) shows an example of the type of the channel. Under the circumstance, we always have $P(o_i/k_i) = 1$ and $P(k_i/o_i) = 1$. As we can see, the channel have the unique output ($o \in O$) for every input ($k \in K$).

@@

\subsubsection{Noiseless Loss Channel}
In reality, noiseless loss channel is more common. The channel capacity of a noiseless loss channel equals to

\begin{equation} \label{chapter:eq:5}
  C(K;O) = \max_{p(x)} I(K;O) = \max_{p(x)} H |O| 
\end{equation}

Figure~\ref{fig:channel}(b) shows an example of the type of the channel. Once the input value is determined, the output value is also determined. Shown in Figure~\ref{fig:channel}(b), the channel has the unique output ($o \in O$) for every input ($k \in K$).

@@

We consider an access-based attack~\cite{DBLP:journals/jce/GeYCH18}, where an adversary can probe one time for each run only after the termination of a deterministic program. The attacker may not know the memory access the victim program accesses directly. Still, he can infer whether some memory addresses are accessed or not at various granularities after \textit{each run} of the victim program. We call the memory access information as the observation ($O$) in the rest of the chapter. It is a typical scenario for most side-channel attacks.  We also assume the attacker can have access to the source code of the victim program ($\beta$). It is a valid assumption since most side-channel attacks target open-source software. Similar to recent work, we assume the attacker has a noise-free observation.  The program ($\beta$) has $K$ as the sensitive input. $K$ is a finite set that consists of every possible key ($k \in K$). The program also takes known messages $M$ as the input.

\begin{figure}[ht]
  \centering
  \includegraphics[width=.8\columnwidth]{./figures/chapter5/channel.pdf}
  \caption{The relationship between the side-channel attack and the channel capacity. For a deterministic program, the side-channel attack can be seen as the communication between the observation ($o \in O$) and the secret ($k \in K$) via a discrete memoryless channel.}
  \label{fig:side_channel}
\end{figure}

With the above definitions, we define the following mapping between $\beta$,
$K$, $M$, and $O$:

\begin{displaymath}
  \beta(K, M) \rightarrow O
\end{displaymath}

Side-channel attacks can be seen as a process that infers the secret ($K$) based on the observations ($O$). As shown in the figure~\ref{fig:side_channel}, we assume that an attacker can observe the memory address of the victim program. He wants to infer the original secret ($k$) based on the observation. We also assume the attacker knows the known message $m \in M$. In the rest of the chapter, we use $\beta(K) \rightarrow O$ to denote a victim program that may have side-channel leakage sites.

@@

To assess the sensitive level of the vulnerability, we use the sampling method to estimate the leakage. Suppose we give the function input from $0, 1, \dots, 7$ and observe the array $T$, we can find each different input has one unique observation. Therefore, the secret can be uniquely determined.

@@

\begin{figure}[h]
  \begin{minipage}{0.40\linewidth}
      \begin{tabular}{rrrr}
        \toprule
        & $a$ & $\lnot a$  & total\\
        \midrule
        $b$   & 255 (255.5) & 253 (252.5) &   508  \\
        $\lnot b$   & 260 (259.5)  & 256 (256.5) &    516   \\
        total &   515 &  509   & 1,024   \\
        \bottomrule
      \end{tabular}\caption*{(a)}
  \end{minipage}
\hspace{-15pt}
\hfill
\hspace{-15pt}
  \begin{minipage}{0.40\linewidth}
      \begin{tabular}{rrrr}
        \toprule
        & $c$ & $\lnot c$  & total\\
        \midrule
        $d$   & 809 (764.8) & 82 (126.2) & 891    \\
        $\lnot d$   & 70 (114.2)  & 63 (18.8)&  133     \\
        total &  879 &  145  & 1,024    \\
        \bottomrule
      \end{tabular}\caption*{(b)}
  \end{minipage}
  \caption{The contingency table for the experiments. The numbers in the bracket are expected values. }\label{chapter5:fig:con_table}
\end{figure}

>>>>

\chapter{Precise Analysis on Multiple-trace Attacks}\label{chapter5}
\section{Introduction}
Side-channel attacks allow attackers to infer sensitive information based on the non-functional behaviors of the computer system. Examples of these non-functional behaviors include acoustics, CPU usages, timing, and EM signals~\cite{agrawal2002side,hund2013practical,halevi2015keyboard,batina2019csi}. In general, if a program has different execution behaviors when it processes various keys, an attacker can infer nontrivial information by observing secret-dependent behaviors. While separate side-channel attacks may have different forms and patterns, a large portion of them share the same fundamental reasons. That is, the memory access pattern is dependent on the original sensitive input.

@@

Even a minor leakage can be severe for cryptography libraries because
the leakage can reduce the encryption algorithm's strength. However,
some attacks also exploit non-cryptography libraries and applications
such as machine learning
applications~\cite{yan2020cache,hong2018security}, graphic
libraries~\cite{wang2019unveiling}, and spell checking
tools~\cite{xu2015controlled}. Unlike side-channel attacks on
cryptography libraries that exploit one vulnerability at a time,
attacks on non-cryptography rely on multiple side-channel leakage
sites to retrieve more information. A program  usually has more than two
leakage sites. However, no existing tools that can estimate the total
effect of multiple leakage sites. Are they independent or have the same root cause? For real-world applications with
thousands of lines, attacker can exploit more than one leakage sites
to retrieve more information. Attackers can guess the values of many temporary
values. These temporary values may contain some knowledge of the
original secret buffer.

Many side-channel leakage detection tools rely on
techniques such as symbolic executions, taint analysis, and abstract interpretations~\cite{203878,182946,Brotzman19Casym,236338}. Due
to the limitations of existing binary analysis techniques, it is hard to apply existing methods to analyze side-channels in non-cryptography
libraries. For many other applications like machine learning, media,
and graphic libraries built on the top architecture-dependent
libraries (e.g., Intel MKL~\cite{wang2014intel}), these tools are not able to
handle them. Moreover, some binary analysis frameworks (e.g.,
Angr~\cite{shoshitaishvili2016state}) are limited in analyzing
floating-point instructions. However, the functionality of these
applications heavily depends on floating-point calculations.

While many existing works can detect side-channel leakages, few
of them can assess the severity levels of these
vulnerabilities~\cite{182946,203878,182946,236338,217537,langley2010ctgrind,xiao2017stacco,wang2019time,Wichelmann:2018:MFF:3274694.3274741,bao2021abacus,236334,wang2021data,sung2018canal}.
Some existing tools~\cite{182946,chattopadhyay2019quantifying, biondi2018scalable,5207642,pasareanu2016multi,AskarovC12,phan2017synthesis} quantify the information leakage by giving an upper
bound estimation of each leakage site.  This is helpful to justify that the software is secure if the reported leakage is zero. However,
these tools cannot report to software developers how severe each leakage site would be because of the over-approximation they apply when they estimate the amount of leakage.  Often the over-approximated estimate is very far away from the amount of real leakage.
For example, CacheAudit~\cite{182946} estimates that a 128-bit
AES encryption can leak more than 128 bits. As a result, even these
tools report that some leakages can leak lots of information, it does not mean the program has truly severe side-channel vulnerabilities.

To solve the above problem, we propose a method to estimate the effect of side-channel leakages automatically. We treat the side-channel attack as a communication system and use the channel capacity to quantify the amount of the information flow from the original secret to an attacker's observation. The method also allows us to combine multiple leakages sites to estimate the overall effect. The attack can be seen as a process that reduces the search space of the original sensitive inputs. We use the side-channel vulnerability to divide the input space. If the patterns can uniquely distinguish the input space, then we think the information is leaked totally. However, in real cases, many of these side-channel leakage patterns are not sufficient to distinguish each leakage site but can still discern some secret inputs. In these cases, we call them partial leakages.

@@

  \item We propose a method that automatically detects and analyzes the
  effect of each side-channel leakage site. Our analysis combines
  information from both the source code and the runtime execution,
  which makes the method more effective in finding leakages. The
  technique can combine multiple leakage sites to retrieve more
  information as well.

@@

\section{Background and Threat Model}
In this chapter, we treat the side-channel attack as a communication system between two parties.
Here we first give an overview of the background knowledge of information theory and present the threat model.

\subsection{Channel Capacity}
Given a discrete random variable $K$,
$k$ is one of the possible
values occurring with the probability $p(k)$. The Shannon entropy $H(K)$ is defined as
\begin{equation}\label{chapter5:eq1}
  H(K) = - \sum_{k {\in} K}p(k)\log_2p(k).
\end{equation}

@@

\begin{equation}\label{chapter5:eq2}
  C(K;O) = \max_{p(x)} I(K;O).
\end{equation}
Here $I(K;O)$ is the mutual information between $K$ and $O$.
\begin{equation} \label{chapter5:eq3}
  I(K;O) = \sum_{k {\in} K}{\sum_{o {\in} O}{p(k, o)\log_2\frac{p(k, o)}{p(k)p(o)}}}
\end{equation}

@@

\subsubsection{Noiseless Lossless Channel}
An information channel is noiseless and lossless if every $k$ corresponds to exactly one unique $o$. The channel capacity of a noiseless lossless channel equals
\begin{equation} \label{chapter5:eq4}
  C(K;O) = \max_{p(k)} I(K;O) = \max_{p(k)} H|O| =\max_{p(k)} H|K|,
\end{equation}
\noindent where $|K|$ represents the number of possible variables in $K$.

Figure~\ref{fig:channel}(a) shows an example of the type of the channel. Under the circumstance, we always have $P(o_i/k_i) = 1$ and $P(k_i/o_i) = 1$. As we can see, the channel has the unique output ($o \in O$) for every input ($k \in K$).

@@

\subsubsection{Noiseless Loss Channel}
In reality, noiseless loss channels are more common. The channel capacity of a noiseless loss channel equals

\begin{equation} \label{chapter:eq:5}
  C(K;O) = \max_{p(x)} I(K;O) = \max_{p(x)} H |O|.
\end{equation}

Figure~\ref{fig:channel}(b) shows an example of this type of channels. Once the input value is determined, the output value is also determined. Shown in Figure~\ref{fig:channel}(b), the channel has the unique output ($o \in O$) for every input ($k \in K$).

@@

We consider an access-based attack~\cite{DBLP:journals/jce/GeYCH18}, where an adversary can probe one time for each run only after the termination of a deterministic program. The attacker may not know the memory access the victim program accesses directly. Still, he can infer whether some memory addresses are accessed or not at various granularities after \textit{each run} of the victim program. We call the memory access information as the observation ($O$) in the rest of the chapter. It is a typical scenario for most side-channel attacks.  We also assume the attacker can have access to the source code of the victim program ($\beta$). It is a valid assumption since most side-channel attacks target open-source software. Similar to recent work, we assume the attacker has a noise-free observation.  The program ($\beta$) has $K$ as the sensitive input, which is a finite set consisted of every possible key ($k \in K$). The program also takes known messages $M$ as the input.

\begin{figure}[ht]
  \centering
  \includegraphics[width=.75\columnwidth]{./figures/chapter5/channel.pdf}
  \caption{The relationship between the side-channel attack and the channel capacity. For a deterministic program, the side-channel attack can be seen as the communication between the observation ($o \in O$) and the secret ($k \in K$) via a discrete memoryless channel.}
  \label{fig:side_channel}
\end{figure}

With the above setting, we define the following mapping between $\beta$,
$K$, $M$, and $O$:

\begin{displaymath}
  \beta(K, M) \rightarrow O.
\end{displaymath}

Side-channel attacks can be seen as a process that infers the secret ($K$) based on the observations ($O$). As shown in Figure~\ref{fig:side_channel}, we assume that an attacker can observe the memory addresses of the victim program. He wants to infer the original secret ($k$) based on the observation. We also assume the attacker knows the known message $m \in M$. In the rest of the chapter, we use $\beta(K) \rightarrow O$ to denote a victim program that may have side-channel leakage sites.

@@

To assess the severity level of the vulnerability, we use the sampling method to estimate the leakage. Suppose we give the function input from $0, 1, \dots, 7$ and observe the array $T$, we can find each different input has one unique observation. Therefore, the secret can be uniquely determined.

@@

\begin{figure}[h]
  \begin{minipage}{0.45\linewidth}
      \begin{tabular}{rrrr}
        \toprule
        & $a$ & $\lnot a$  & total\\
        \midrule
        $b$   & 255 (255.5) & 253 (252.5) &   508  \\
        $\lnot b$   & 260 (259.5)  & 256 (256.5) &    516   \\
        total &   515 &  509   & 1,024   \\
        \bottomrule
      \end{tabular}\caption*{(a)}
  \end{minipage}
%\hspace{-15pt}
\hfill
%\hspace{-15pt}
  \begin{minipage}{0.45\linewidth}
      \begin{tabular}{rrrr}
        \toprule
        & $c$ & $\lnot c$  & total\\
        \midrule
        $d$   & 809 (764.8) & 82 (126.2) & 891    \\
        $\lnot d$   & 70 (114.2)  & 63 (18.8)&  133     \\
        total &  879 &  145  & 1,024    \\
        \bottomrule
      \end{tabular}\caption*{(b)}
  \end{minipage}
  \caption{The contingency table for the experiments. The numbers in the bracket are expected values. }\label{chapter5:fig:con_table}
\end{figure}
